---
title: "haplo-stats"
description: |
  impute high resolution (four digit) alleles; using haplo-stats algorithm
output:
  rmarkdown::html_vignette: default
vignette: |
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{hla-stats}
  %\usepackage[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## work in progress

```{r}
library(hlaR)
library(tidyverse)
library(readxl)
```

## import haplostats reference data

Provide references and links.  Describe data and approach

```{r}
haplo_stats <- read_xlsx("data/A~C~B~DRB3-4-5~DRB1~DQB1.xlsx")

```

    
## Brief exploration of Haplostats 

most common haplotypes by geography and ethnicity



```{r}

hla_common <- haplo_stats %>% 
  select(A:DQB1, AAFA_rank, EURCAU_rank, EURCAU_rank, MSWHIS_rank, NCHI_rank) %>% 
  mutate(across(contains("rank"), as.integer)) %>% 
  filter(AAFA_rank <= 10 | EURCAU_rank <= 10 | MSWHIS_rank<= 10 | NCHI_rank<= 10)


```

##  create some sample subjects

For each group create hypothetical subject by pairing the haplo ranks 1/2, 3/4, 5/6 for each ethnicity/geo
3 EURCAU - 
3 AAFA - ranks 1/2, 3/4, 5/6
3 HIS
3 NCHI


## Describe function

## Run sample data

## 

